---
title: "a03"
author: Jorge coronado
---
# Introduction

This analysis shows the catch pattern by region for different species. This information is useful for fisheries management, conservation and making decisions based on data.

# Data

The data comes from an ecological repository and it was read directly from the source.

# Analysis

This analysis first removes irrelevant columns and fixes the format of the Chinook column. Then, the data set is transformed into long format, the catch values are multiplied by 1000 to adjust the units, and the mean catch is calculated by region. Finally, the regions are ordered from highest to lowest mean catch.

## Setup

```{r}
library(dplyr)
library(tidyr)
library(readr)
```

```{r}
# read the csv file from the url using libcurl.
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"))
```

```{r}
mean_region <- catch_original %>% 
  select(-All, -notesRegCode) %>% # remove not needed columns for calculating means
  mutate(Chinook = if_else(Chinook == "I", "1", Chinook)) %>% # replace "I" values for "1" in the Chinook column
  mutate(Chinook = as.numeric(Chinook)) %>% # convert the chinook column from character to numeric
  pivot_longer(-c(Region, Year), # reshape data from wide format to long format, keeping Region and Year
               names_to = "species", # new column containing species names
               values_to = "catch") %>% # new column containing catch values
  mutate(catch = catch*1000) %>% # multiply catch values by 1000
  group_by(Region) %>%  # group data by region 
  summarize(mean_catch = mean(catch)) %>% # calculate the mean of catch by region
  arrange(desc(mean_catch)) # sort out in decreasing order the values of mean catch

head(mean_region) # display the first few rows of mean_region
```
# Discussion

Regions with higher mean_catch stand out as areas with greater average catch. However, it is important to check whether these averages are influenced by a few extreme years. Before making casual interpretations, it is recommended to review values (NA), check for outliers, confirm the units, and consider using medians or interquartile ranges to complement the mean...